Contextual Combinatorial Cascading Bandits

### 2015-12-10 contextual combinatorial cascading bandits.tex
update main proof and put lemma 10 after lemma 12

@@

\begin{lemma}
\label{lem:SumXiEstimateInDet}
Let $x_i \in \RR^{d \times 1}, 1 \leq i \leq n$. Then we have
$$
\det\left(I + \sum_{i=1}^n x_i x_i^{\top}\right) \geq 1 + \sum_{i=1}^n \norm{x_i}_2^2.
$$
\end{lemma}
\begin{proof}
Denote the eigenvalues of $I + \sum_{i=1}^n x_i x_i^{\top}$ by $1+\alpha_1,...,1+\alpha_d$ with $\alpha_j \geq 0$, $1\leq j\leq d$. Then
\begin{align*}
&\det(I + \sum_{i=1}^n x_i x_i^{\top})= \prod_{j=1}^d (1 + \alpha_j)\\
&\geq 1 +\sum_{j=1}^d \alpha_j =1-d + \sum_{i=1}^d (1+\alpha_i) \\
&=1-d + \trace(I + \sum_{i=1}^n x_i x_i^{\top})= 1-d + d + \sum_{i=1}^n \norm{x_i}_2^2\\
&=1 + \sum_{i=1}^n \norm{x_i}_2^2.
\end{align*}
\end{proof}

\begin{lemma}
If $\lambda \geq C_\gamma$, then
$$
\sum_{s=1}^t \sum_{k=1}^{\bO_s} \norm{\gamma_k x_{s,\ba_{k}^s}}_{\bV_{s-1}^{-1}}^2 \leq 2\ln \left(\frac{\det(\bV_t)}{\lambda^d} \right).
$$
\end{lemma}
\begin{proof}
\begin{align*}
&\det(\bV_t) = \det(\bV_{t-1} + \sum_{k=1}^{\bO_t} (\gamma_k X_{t,\ba_k^{t}})(\gamma_k X_{t, \ba_k^{t}}^{\top}))\\
&=\det(\bV_{t-1})\\
&~~~\cdot \det(I + \sum_{k=1}^{\bO_t} \gamma_k \bV_{t-1}^{-1/2}x_{t,\ba_{k}^{t}} (\gamma_k \bV_{t-1}^{-1/2}x_{t,\ba_{k}^{t}})^{\top})\\
&\geq \det(\bV_{t-1}) (1 + \sum_{k=1}^{\bO_t} \norm{\gamma_k x_{t,\ba_k^t}}_{\bV_{t-1}^{-1}}^2)\\
&\geq \det(\lambda I)\prod_{s=1}^{t}(1 + \sum_{k=1}^{\bO_s} \norm{\gamma_k x_{s,\ba_k^s}}_{\bV_{s-1}^{-1}}^2)
\end{align*}
Because
$$
\norm{\gamma_k x_{s,\ba_k^s}}_{\bV_{s-1}^{-1}}^2 \leq \norm{\gamma_k x_{s,\ba_k^s}}_2^2/\lambda_{\min}(\bV_{s-1}) \leq \gamma_k^2 /\lambda,
$$
we have 
$$
\sum_{k=1}^{\bO_s} \norm{\gamma_k x_{s,\ba_k^s}}_{\bV_{s-1}^{-1}}^2 \leq \frac{1}{\lambda} \sum_{k=1}^{K} \gamma_k^2 = C_\gamma /\lambda \leq 1
$$
Using the fact that $ 2\ln(1+u) \geq u$ for any $u \in [0,1]$, we get
\begin{align*}
&\sum_{s=1}^t \sum_{k=1}^{\bO_s}\norm{\gamma_k x_{s,\ba_{k}^s}}_{\bV_{s-1}^{-1}}^2 \\
&\leq 2\sum_{s=1}^t\ln(1 + \sum_{k=1}^{\bO_s} \norm{\gamma_k x_{s,\ba_k^s}}_{\bV_{s-1}^{-1}}^2)\\
&\leq 2(\ln(\det(\bV_t)) - \ln(\det(\lambda I))) \\
&= 2 \ln \left(\frac{\det(\bV_t)}{\det(\lambda I)} \right)
\end{align*}
\end{proof}

@@

For the disjunctive case,
\begin{equation}
\begin{split}
&R(T) = \EE[\sum_{t=1}^{T} \EE_{t}[R(t, \bA_t)] ]\\
&\leq \EE\left[\sum_{t=1}^{T} \frac{1}{p^*} \EE_t \left[\bbeta_{t-1}(\delta) \sum_{k=1}^{\bO_t}\norm{\gamma_k x_{t,\ba_k^t}}_{\bV_{t-1}^{-1}}\right] \right]\\
&\leq \EE \left[\frac{1}{p^*} \bbeta_T(\delta) \sqrt{\left(\sum_{t=1}^{T} \bO_t\right) \left( \sum_{t=1}^{T} \sum_{k=1}^{\bO_t}\norm{\gamma_k x_{t,\ba_k^t}}_{\bV_{t-1}^{-1}}^2 \right)}
	\right ]  \\
&\leq \EE\left[\frac{1}{p^*} \left( \sqrt{ \ln \left( \frac{\det(\bV_T)}{\lambda^d \delta^2} \right)} + \sqrt{\lambda} \right) \right. \\
& \left. \qquad \qquad \cdot \sqrt{TK \cdot 2\ln \left(\frac{\det(\bV_T)}{\lambda^d} \right)} \right]\\
&\leq \frac{\sqrt{2}}{p^*} \left(\sqrt{ d \ln \left( \frac{1 + C_\gamma T/(\lambda d)}{\delta^2} \right) } + \sqrt{\lambda} \right)\\
&\qquad \qquad \cdot \sqrt{TKd\ln(1 + C_\gamma T/(\lambda d))} ]
\end{split}
\end{equation}
The first inequality is by Lemma \ref{lem:DeltaEsimateWithP*}. The second inequality is because $\bbeta_{t}(\delta)$ is increasing with respect to $t$, derived by definition of $\bbeta_t(\delta)$ and Lemma \ref{lem:detVt}, and mean inequality. The third inequality is because of the definition of $\beta_t(\delta)$ and Lemma \ref{lem:SumXiEstimateInDet}. And the last inequality is because estimate of $\det(\bV_t)$ by Lemma \ref{lem:detVt}.

Similarly, for the conjunctive case, 
\begin{equation}
\begin{split}
&R(T) = \EE[ \sum_{t=1}^{T} \EE_{t}[R_{\wedge}(t, \bA_t)] ]\\
&\leq \EE \left[\sum_{t=1}^{T} \frac{8}{f_{\wedge}^{\ast}} \EE_t\left[\bbeta_{t-1}(\delta)\sum_{k=1}^{\bO_t}\norm{\gamma_k x_{t,\ba_k^t}}_{\bV_{t-1}^{-1}} \right] \right] \\
&\leq \EE\left[\frac{8}{f_{\wedge}^{\ast}} \bbeta_{T}(\delta) \sum_{t=1}^{T} \sum_{k=1}^{\bO_t}\norm{\gamma_k x_{t,\ba_k^t}}_{\bV_{t-1}^{-1}} \right]\\
&\leq \EE\left[\frac{8}{f_{\wedge}^{\ast}} \bbeta_{T}(\delta) \sqrt{ \left(\sum_{t=1}^{T} \bO_t \right) \EE_t \left[\sum_{t=1}^{T} \sum_{k=1}^{\bO_t}\norm{\gamma_k x_{t,\ba_k^t}}_{\bV_{t-1}^{-1}}^2 \right]} \right]\\
&\leq \EE \left[\frac{\sqrt{128}}{f_{\wedge}^{\ast}} \left(\sqrt{\ln\left(\frac{\det(\bV_T)}{\lambda^d \delta^2}\right) } + \sqrt{\lambda} \right) \right. \\
&\qquad \qquad \left. \cdot \sqrt{TK \cdot 2\ln \left(\frac{\det(\bV_T)}{\lambda^d} \right)} \right]\\
&\leq \frac{\sqrt{128}}{f_{\wedge}^{\ast}} \left(\sqrt{d \ln \left( \frac{1 + C_\gamma T/(\lambda d)}{\delta^2}\right) } + \sqrt{\lambda} \right)\\
&\qquad \qquad \cdot \sqrt{TKd\ln(1 + C_\gamma T/(\lambda d))} ]
\end{split}
\end{equation}
\end{proof}

The first inequality is by Lemma \ref{lem:DeltaEsimateWithP*}. The second inequality is because $\bbeta_{t}(\delta)$ is increasing with respect to $t$, derived by definition of $\bbeta_t(\delta)$ and Lemma \ref{lem:detVt}. The third inequality is because of mean inequality. The forth inequality is by the definition of $\beta_t(\delta)$ and Lemma \ref{lem:SumXiEstimateInDet}. And the last inequality is because estimate of $\det(\bV_t)$ by Lemma \ref{lem:detVt}.

>>>>

From the above lemma, we can bound the regret at time $t$ in terms of the observed base arms, which we would like to estimate in the following. Before that, we need a technical lemma.

\begin{lemma}
\label{lem:detTech}
Let $x_i \in \RR^{d \times 1}, 1 \leq i \leq n$. Then we have
$$
\det\left(I + \sum_{i=1}^n x_i x_i^{\top}\right) \geq 1 + \sum_{i=1}^n \norm{x_i}_2^2.
$$
\end{lemma}
\begin{proof}
Denote the eigenvalues of $I + \sum_{i=1}^n x_i x_i^{\top}$ by $1+\alpha_1,...,1+\alpha_d$ with $\alpha_j \geq 0$, $1\leq j\leq d$. Then
\begin{align*}
&\det(I + \sum_{i=1}^n x_i x_i^{\top})= \prod_{j=1}^d (1 + \alpha_j)\\
&\geq 1 +\sum_{j=1}^d \alpha_j =1-d + \sum_{i=1}^d (1+\alpha_i) \\
&=1-d + \trace(I + \sum_{i=1}^n x_i x_i^{\top})= 1-d + d + \sum_{i=1}^n \norm{x_i}_2^2\\
&=1 + \sum_{i=1}^n \norm{x_i}_2^2.
\end{align*}
\end{proof}

Now we can give the following estimate about all observed base arms.

\begin{lemma}
\label{lem:SumXiEstimateInDet}
If $\lambda \geq C_\gamma$, then
$$
\sum_{s=1}^t \sum_{k=1}^{\bO_s} \norm{\gamma_k x_{s,\ba_{k}^s}}_{\bV_{s-1}^{-1}}^2 \leq 2\ln \left(\frac{\det(\bV_t)}{\lambda^d} \right).
$$
\end{lemma}
\begin{proof}
\begin{align}
&\det(\bV_t) = \det(\bV_{t-1} + \sum_{k=1}^{\bO_t} (\gamma_k X_{t,\ba_k^{t}})(\gamma_k X_{t, \ba_k^{t}}^{\top})) \notag\\
&=\det(\bV_{t-1}) \notag \\
&~~~\cdot \det(I + \sum_{k=1}^{\bO_t} \gamma_k \bV_{t-1}^{-1/2}x_{t,\ba_{k}^{t}} (\gamma_k \bV_{t-1}^{-1/2}x_{t,\ba_{k}^{t}})^{\top}) \notag\\
&\geq \det(\bV_{t-1}) (1 + \sum_{k=1}^{\bO_t} \norm{\gamma_k x_{t,\ba_k^t}}_{\bV_{t-1}^{-1}}^2) \label{eq:detGeq} \\
&\geq \det(\lambda I)\prod_{s=1}^{t}(1 + \sum_{k=1}^{\bO_s} \norm{\gamma_k x_{s,\ba_k^s}}_{\bV_{s-1}^{-1}}^2), \notag
\end{align}
where inequality \eqref{eq:detGeq} is because of Lemma \ref{lem:detTech}.

Because
$$
\norm{\gamma_k x_{s,\ba_k^s}}_{\bV_{s-1}^{-1}}^2 \leq \norm{\gamma_k x_{s,\ba_k^s}}_2^2/\lambda_{\min}(\bV_{s-1}) \leq \gamma_k^2 /\lambda,
$$
we have 
$$
\sum_{k=1}^{\bO_s} \norm{\gamma_k x_{s,\ba_k^s}}_{\bV_{s-1}^{-1}}^2 \leq \frac{1}{\lambda} \sum_{k=1}^{K} \gamma_k^2 = C_\gamma /\lambda \leq 1
$$
Using the fact that $ 2\ln(1+u) \geq u$ for any $u \in [0,1]$, we get
\begin{align*}
&\sum_{s=1}^t \sum_{k=1}^{\bO_s}\norm{\gamma_k x_{s,\ba_{k}^s}}_{\bV_{s-1}^{-1}}^2 \\
&\leq 2\sum_{s=1}^t\ln(1 + \sum_{k=1}^{\bO_s} \norm{\gamma_k x_{s,\ba_k^s}}_{\bV_{s-1}^{-1}}^2)\\
&\leq 2(\ln(\det(\bV_t)) - \ln(\det(\lambda I))) \\
&= 2 \ln \left(\frac{\det(\bV_t)}{\det(\lambda I)} \right)
\end{align*}
\end{proof}

By the above lemma, we need to estimate $\det(\bV_t)$ to bound regret.

@@

@@

For the disjunctive case,
\begin{align}
&R(T) = \EE[\sum_{t=1}^{T} \EE_{t}[R(t, \bA_t)] ] \notag\\
&\leq \EE\left[\sum_{t=1}^{T} \frac{1}{p^*} \EE_t \left[\bbeta_{t-1}(\delta) \sum_{k=1}^{\bO_t}\norm{\gamma_k x_{t,\ba_k^t}}_{\bV_{t-1}^{-1}}\right] \right] \label{eq:pfMdeltaEst}\\
&\leq \EE\left[\sum_{t=1}^{T} \frac{1}{p^*} \bbeta_{T}(\delta) \EE_t \left[ \sum_{k=1}^{\bO_t}\norm{\gamma_k x_{t,\ba_k^t}}_{\bV_{t-1}^{-1}}\right] \right] \label{eq:pfMbeta}\\
&\leq \EE \left[\frac{1}{p^*} \bbeta_T(\delta) \sqrt{\left(\sum_{t=1}^{T} \bO_t\right) \left( \sum_{t=1}^{T} \sum_{k=1}^{\bO_t}\norm{\gamma_k x_{t,\ba_k^t}}_{\bV_{t-1}^{-1}}^2 \right)}
	\right ]  \label{eq:pfMmeanEq}\\
&\leq \EE\left[\frac{1}{p^*} \left( \sqrt{ \ln \left( \frac{\det(\bV_T)}{\lambda^d \delta^2} \right)} + \sqrt{\lambda} \right) \right. \notag\\
& \left. \qquad \qquad \cdot \sqrt{TK \cdot 2\ln \left(\frac{\det(\bV_T)}{\lambda^d} \right)} \right] \label{eq:pfMsumXi}\\
&\leq \frac{\sqrt{2}}{p^*} \left(\sqrt{ d \ln \left( \frac{1 + C_\gamma T/(\lambda d)}{\delta^2} \right) } + \sqrt{\lambda} \right) \notag\\
&\qquad \qquad \cdot \sqrt{TKd\ln(1 + C_\gamma T/(\lambda d))} ] \label{eq:pfMdetEst}
\end{align}
The inequality \eqref{eq:pfMdeltaEst} is by Lemma \ref{lem:DeltaEsimateWithP*}. The second inequality \eqref{eq:pfMbeta} is because $\bbeta_{t}(\delta)$ is increasing with respect to $t$, derived by definition of $\bbeta_t(\delta)$ and Lemma \ref{lem:detVt}. The third inequality \eqref{eq:pfMmeanEq} is because of mean inequality. The fourth inequality \eqref{eq:pfMsumXi} is because of the definition of $\beta_t(\delta)$ and Lemma \ref{lem:SumXiEstimateInDet}. And the last inequality \eqref{eq:pfMdetEst} is because estimate of $\det(\bV_t)$ by Lemma \ref{lem:detVt}.

Similarly, for the conjunctive case, 
\begin{align}
&R(T) = \EE[ \sum_{t=1}^{T} \EE_{t}[R_{\wedge}(t, \bA_t)] ] \notag\\
&\leq \EE \left[\sum_{t=1}^{T} \frac{8}{f_{\wedge}^{\ast}} \EE_t\left[\bbeta_{t-1}(\delta)\sum_{k=1}^{\bO_t}\norm{\gamma_k x_{t,\ba_k^t}}_{\bV_{t-1}^{-1}} \right] \right] \label{eq:pfMCdeltaEst}\\
&\leq \EE\left[\frac{8}{f_{\wedge}^{\ast}} \bbeta_{T}(\delta) \sum_{t=1}^{T} \sum_{k=1}^{\bO_t}\norm{\gamma_k x_{t,\ba_k^t}}_{\bV_{t-1}^{-1}} \right] \label{eq:pfMCbeta}\\
&\leq \EE\left[\frac{8}{f_{\wedge}^{\ast}} \bbeta_{T}(\delta) \sqrt{ \left(\sum_{t=1}^{T} \bO_t \right) \EE_t \left[\sum_{t=1}^{T} \sum_{k=1}^{\bO_t}\norm{\gamma_k x_{t,\ba_k^t}}_{\bV_{t-1}^{-1}}^2 \right]} \right] \label{eq:pfMCmeaneq}\\
&\leq \EE \left[\frac{\sqrt{128}}{f_{\wedge}^{\ast}} \left(\sqrt{\ln\left(\frac{\det(\bV_T)}{\lambda^d \delta^2}\right) } + \sqrt{\lambda} \right) \right. \notag\\
&\qquad \qquad \left. \cdot \sqrt{TK \cdot 2\ln \left(\frac{\det(\bV_T)}{\lambda^d} \right)} \right] \label{eq:pfMCsumXi}\\
&\leq \frac{\sqrt{128}}{f_{\wedge}^{\ast}} \left(\sqrt{d \ln \left( \frac{1 + C_\gamma T/(\lambda d)}{\delta^2}\right) } + \sqrt{\lambda} \right) \notag\\
&\qquad \qquad \cdot \sqrt{TKd\ln(1 + C_\gamma T/(\lambda d))} ] \label{eq:pfMCdetEst}
\end{align}
\end{proof}

The first inequality \eqref{eq:pfMCdeltaEst} is by Lemma \ref{lem:DeltaEsimateWithP*}. The second inequality \eqref{eq:pfMCbeta} is because $\bbeta_{t}(\delta)$ is increasing with respect to $t$, derived by definition of $\bbeta_t(\delta)$ and Lemma \ref{lem:detVt}. The third inequality \eqref{eq:pfMCmeaneq} is because of mean inequality. The forth inequality \eqref{eq:pfMCsumXi} is by the definition of $\beta_t(\delta)$ and Lemma \ref{lem:SumXiEstimateInDet}. And the last inequality \eqref{eq:pfMCdetEst} is because estimate of $\det(\bV_t)$ by Lemma \ref{lem:detVt}.
